Systematic studies of correlations between different order flow harmonics in $\PbPb$ collisions at $\snn = 2.76$~TeV

### 2017-08-09 longSC.tex
final to PC

\section{Introduction}
\label{sec:intro}
The main emphasis of the ultra-relativistic heavy-ion collision programs at the Relativistic Heavy Ion Collider (RHIC) and the Large Hadron Collider (LHC) is to study the deconfined phase of strongly interacting nuclear matter, the Quark-Gluon Plasma (QGP). 
%This matter exhibits strong anisotropic flow in the plane transverse to the beam direction, which is driven by anisotropic pressure gradients, resulting in more particles emitted in the direction of the largest gradients.
The matter produced in a heavy-ion collision exhibits strong collective radial expansion. Due to anisotropic pressure gradients in the plane transverse to the beam direction, more particles emitted in the direction of the largest gradients result in anisotropic transverse flow.
The large elliptic flow discovered at RHIC energies~\cite{Ackermann:2000tr} is also observed at LHC energies~\cite{Aamodt:2010pa,Adam:2016izf,Aad:2014eoa}. The measurements are well described by calculations utilizing viscous hydrodynamics~\cite{Romatschke:2007mq,Shen:2011eg,Schenke:2011zz,Bozek:2012qs,Gale:2012rq,Hirano:2010je}.
%and microscopic transport models ~\cite{Xu:2007jv}.  
These calculations also demonstrated that the shear viscosity to the entropy density ratio ($\eta/s$) of the QGP in heavy-ion collisions at RHIC and LHC energies is close to a universal lower bound $1/4\pi$~\cite{Kovtun:2004de}.

@@

\section{Data Analysis}
\label{sec:experiment}
Data recorded by ALICE in $\PbPb$ collisions at $\snn=2.76$~TeV during the 2010 heavy-ion run at the
LHC is used for this analysis. Detailed descriptions of the ALICE
detector can be found
in~\cite{Aamodt:2008zz,Carminati:2004fp,Alessandro:2006yt}. The Time
Projection Chamber (TPC) was used to reconstruct charged particle
tracks and measure their momenta with full azimuthal coverage in the
pseudorapidity range $|\eta|<0.8$. Two scintillator
arrays (V0A and V0C) which cover the pseudorapidity  ranges $-3.7<\eta<-1.7$
and $2.8<\eta<5.1$ were used for triggering and the determination of
centrality~\cite{Aamodt:2010cz}. The trigger
conditions and the event selection criteria are identical to those
described in~\cite{Aamodt:2010pa, Aamodt:2010cz}.
Approximately $10^7$ minimum-bias $\PbPb$ events with
a reconstructed primary vertex within $\pm 10$ cm from the nominal
interaction point along the beam direction are selected. Only charged particles reconstructed in the TPC in $|\eta|<0.8$
and $0.2<p_{\rm T}<5$~GeV/$c$ were included in the analysis. The charged track quality cuts
described in~\cite{Aamodt:2010pa} were applied to minimize
contamination from secondary charged particles and fake tracks.
The track reconstruction efficiency and contamination
were estimated from HIJING Monte Carlo
simulations~\cite{Wang:1991hta} combined with a GEANT3~\cite{Brun:1994aa} detector model and were found to be independent of
the collision centrality. The reconstruction efficiency increases with transverse momenta from
70\% to 80\% for particles with $0.2<p_{\rm T}<1$~GeV/$c$ and remains
constant at $(80 \pm 5)$\% for $p_{\rm T}>1$~GeV/$c$. The estimated
contamination by secondary charged particles from weak decays and
photon conversions is less than 6\% at $p_{\rm T}=0.2$~GeV/$c$ and falls
below 1\% for $p_{\rm T}>1$~GeV/$c$.
The $p_{\rm T}$ cut-off of 0.2~GeV/$c$ reduces event-by-event biases due to small reconstruction efficiency 
at lower $p_{\rm T}$, while the high $p_{\rm T}$ cut-off of 5~GeV/$c$ reduces the effects of jets on the measured correlations. 
Reconstructed TPC tracks are required to have at least 70 space points (out of a maximum of 159). 
Only tracks with a transverse distance of closest approach to the primary vertex less than 3 mm, both in the longitudinal and transverse directions, are accepted. This reduces the contamination from secondary tracks produced in the detector material, particles from weak decays, etc. Tracks with kinks (i.e. tracks that appear to change direction due to multiple scattering or $K^{\pm}$ decays) were rejected.

@@

\subsection{Centrality Dependence of SC($m$,$n$) and NSC($m$,$n$)}
\label{sec:theory_allorder}
Three different models are compared with the experimental results presented in Sec.~\ref{sec:results}. 
We compare to event-by-event EKRT+viscous hydrodynamic predictions with various parameterizations of the temperature dependence of $\eta/s(T)$; these were shown in Fig.~2 of Ref.~\cite{ALICE:2016kpq}.
In this previous work it was demonstrated that NSC(3,2) is sensitive mainly to the initial conditions, while NSC(4,2) is sensitive to both the initial conditions and the system properties, which is consistent with the predictions from~\cite{Niemi:2012aj}.
However, the sign of NSC(3,2) is negative in the data in 0--10\% central collisions while it is positive in the models where the anisotropies originate mainly from fluctuations.
This observation helps us to better understand the fluctuations in initial energy density.
The NSC(4,2) observable shows better sensitivity for different $\eta/s(T)$ parameterizations but the model cannot describe either the centrality dependence or the absolute values. This observed discrepancy between data and theoretical predictions indicates that the current understanding of initial conditions in models of heavy-ion collisions needs to be revisited to further constrain $\eta/s(T)$.
%, considering the difficulties in separating the role of $\eta/s$ from the initial conditions to the final state particle anisotropies~\cite{Romatschke:2007mq,Shen:2011zc}.
The measurement of SC$(m,n)$ and NSC$(m,n)$ can provide new constraints for the detailed modeling of fluctuating initial conditions.
%Better constraints on the initial state conditions will certainly reduce the uncertainties in determining $\eta/s(T)$.

@@

The comparison to the VISH2+1 calculation~\cite{Zhu:2016puf} is shown in Fig.~\ref{fig:Figure_4}.  All calculations with large $\eta/s$ regardless of the initial conditions ($\eta/s=0.2$ for MC-KLN and MC-Glauber initial conditions and $\eta/s=0.16$ for AMPT initial conditions) fail to capture the centrality dependence of the SC($m$,$n$) observables of all orders, shown in the left panels in Fig.~\ref{fig:Figure_4}.
Among the calculations with small $\eta/s$ ($\eta/s=0.08$), the one with the AMPT initial conditions describes the data better than the ones with other initial conditions for all SC($m$,$n$) observables measured, but it cannot describe the data quantitively for most of the centrality ranges.
Similar to the event-by-event EKRT+viscous hydrodynamic calculations~\cite{Niemi:2015qia}, the sign of the normalized NSC(3,2) in the model calculations in Fig.~\ref{fig:Figure_4} is opposite to that in data in 5--10\% central collisions. 
%NSC(3,2) and NSC(4,3) are not very sensitive to the initial conditions or $\eta/s$ parameterization used in the model, and are not described quantitatively.
The comparisons of the measured NSC(3,2) and NSC(4,3) to VISH2+1 model calculations suggest that they are not very sensitive to the initial conditions or $\eta/s$ parameterization, though the model does not describe them quantitatively.

However, NSC(4,2) is sensitive both to the initial conditions and the $\eta/s$ parameterizations used in the models.
Even though NSC(4,2) favors both AMPT initial conditions with $\eta/s=0.08$ and MC-Glauber initial conditions with $\eta/s=0.20$,
SC(4,2) can only be described by models with smaller $\eta/s$. Hence the calculation with large $\eta/s=0.20$ is ruled out. We conclude that $\eta/s$ should be small and that AMPT initial conditions are favored by the data.
The NSC(5,2) and NSC(5,3) observables are quite sensitive to both the initial conditions and the $\eta/s$ parameterizations.
Similar to the hydrodynamic calculations mentioned above~\cite{Niemi:2015qia}, the sign of NSC(4,3) in these models is opposite to its sign in the data in 0--10\% central collisions. 
The SC(4,3) results are clearly favored by smaller $\eta/s$ values but NSC(4,3) cannot be described by these models quantitively.

@@

The SC($m$,$n$) and NSC($m$,$n$) observables calculated from AMPT simulations are compared with data in Fig.~\ref{fig:Figure_5}.
For SC(3,2), the calculation with the default AMPT settings is closest to the data, but none of the AMPT configurations can describe the data fully. 
The same default calculation can describe the sign and magnitude of NSC(3,2) while the hydrodynamic calculations fail to describe either of them in the most central collisions.
%Interestingly, the string melting AMPT configuration cannot reproduce the data, and the resulting correlations are weaker than in data.
The third version based on the string melting configuration without the hadronic rescattering phase is also shown.
The hadronic rescattering stage makes both SC(3,2) and NSC(3,2) stronger in the string melting AMPT model but not enough to describe the data.
Further investigations proved why the default AMPT model can describe NSC(3,2) fairly well but underestimates SC(3,2). By taking the differences in the individual flow harmonics ($v_2$ and $v_3$) between the model and data into account, it was possible to recover the difference in SC(3,2) between the data and the model. The discrepancy in SC(3,2) can be explained by the overestimated individual $v_n$ values as reported in \cite{Adam:2016nfo} in all centrality ranges.

@@

We list here the important findings from the model comparisons to the centrality dependence of SC($m$,$n$) and NSC($m$,$n$):
\begin{enumerate}[(i)]
	\item The NSC(3,2) observable is sensitive mainly to the initial conditions, while the other observables are sensitive to both the initial conditions and the temperature dependence of $\eta/s$.
	\item The correlation strength between $v_3$ and $v_2$ and between $v_4$ and $v_3$ (SC(3,2), SC(4,2), NSC(3,2) and NSC(4,3)) is significantly underestimated in hydrodynamic model calculations in mid-central collisions.
	\item The sign of NSC(3,2) in 0--10\% central collisions was found to be different in the data and the hydrodynamic model calculations while the default AMPT model can reproduce it.
	\item All the VISH2+1 model calculations with large $\eta/s$ fail to capture the centrality dependence of the correlations regardless of the initial conditions.
	\item Among the VISH2+1 model calculations with small $\eta/s$ ($\eta/s=0.08$), the one with the AMPT initial conditions captures the data qualitatively, but not quantitively for most of the centrality ranges.
	\item The default AMPT model can describe the normalized symmetric cumulants (NSC($m$,$n$)) quantitively for most centralities while the string melting AMPT model fails to describe them.
	%\item The hadronic rescattering in AMPT model makes the correlation strength (NSC) larger.
	\item A hierarchy NSC(5,3) $>$ NSC(4,2) $>$ NSC(5,2) holds for centrality ranges $>~20\%$ within the errors. This hierarchy is reproduced well both by hydrodynamic and AMPT model calculations.

@@

We have found that fluctuations of SC(3,2) and SC(4,3) are anti-correlated in all centralities while fluctuations of SC(4,2), SC(5,2) and SC(5,3) are correlated for all centralities. 
These measurements were compared to various hydrodynamic model calculations with different initial conditions as well as different parameterizations of the temperature dependence of $\eta/s$.
It is found that the different order harmonic correlations have different sensitivities to the initial conditions and the system properties. Therefore they have discriminating power in separating the effects of $\eta/s$  from the initial conditions on the final state particle anisotropies.
The sign of SC(3,2) in 0--10\% central collisions was found to be different between the data and hydrodynamic model calculations.
In the most central collisions the anisotropies originate mainly from fluctuations, where the initial ellipsoidal geometry which dominates in mid-central collisions plays little role. This observation might help to understand the details of the fluctuations in the initial stage. 
The comparisons to VISH2+1 calculations show that all the models with large $\eta/s$, regardless of the initial conditions, fail to capture the centrality dependence of higher order correlations. 
Based on the tested model parameters, the data favors small $\eta/s$ and the AMPT initial conditions.

>>>>

\section{Introduction}
\label{sec:intro}
The main emphasis of the ultra-relativistic heavy-ion collision programs at the Relativistic Heavy Ion Collider (RHIC) and the Large Hadron Collider (LHC) is to study the deconfined phase of strongly interacting nuclear matter, the Quark-Gluon Plasma (QGP). 
%This matter exhibits strong anisotropic flow in the plane transverse to the beam direction, which is driven by anisotropic pressure gradients, resulting in more particles emitted in the direction of the largest gradients.
%The matter produced in a heavy-ion collision exhibits strong collective radial expansion. Due to anisotropic pressure gradients in the plane transverse to the beam direction, more particles emitted in the direction of the largest gradients result in anisotropic transverse flow.
The matter produced in a heavy-ion collision exhibits strong collective radial expansion. 
Geometrical anisotropy of the almond shaped overlap region of the colliding nuclei causes larger pressure gradients into short direction of the almond,
that results in anisotropic transverse flow in the momentum space through interactions of the matter constituents.
The large elliptic flow discovered at RHIC energies~\cite{Ackermann:2000tr} is also observed at LHC energies~\cite{Aamodt:2010pa,Adam:2016izf,Aad:2014eoa}. The measurements are well described by calculations utilizing viscous hydrodynamics~\cite{Romatschke:2007mq,Shen:2011eg,Schenke:2011zz,Bozek:2012qs,Gale:2012rq,Hirano:2010je}.
%and microscopic transport models ~\cite{Xu:2007jv}.  
These calculations also demonstrated that the shear viscosity to the entropy density ratio ($\eta/s$) of the QGP in heavy-ion collisions at RHIC and LHC energies is close to a universal lower bound $1/4\pi$~\cite{Kovtun:2004de}.

@@

\section{Data Analysis}
\label{sec:experiment}
The data samples analyzed in this article were recorded by ALICE during the 2010 heavy-ion run at the LHC in $\PbPb$ collisions at a centre-of-mass energy $\snn=2.76$~TeV.
Detailed descriptions of the ALICE detector can be found
in~\cite{Aamodt:2008zz,Carminati:2004fp,Alessandro:2006yt}. The Time
Projection Chamber (TPC) was used to reconstruct charged particle
tracks and measure their momenta with full azimuthal coverage in the
pseudorapidity range $|\eta|<0.8$. Two scintillator
arrays (V0A and V0C) which cover the pseudorapidity  ranges $-3.7<\eta<-1.7$
and $2.8<\eta<5.1$ were used for triggering and the determination of
centrality~\cite{Aamodt:2010cz}. The trigger
conditions and the event selection criteria are identical to those
described in~\cite{Aamodt:2010pa, Aamodt:2010cz}.
Approximately $10^7$ minimum-bias $\PbPb$ events with
a reconstructed primary vertex within $\pm 10$ cm from the nominal
interaction point along the beam direction are selected. Only charged particles reconstructed in the TPC in $|\eta|<0.8$
and $0.2<p_{\rm T}<5$~GeV/$c$ were included in the analysis. The charged track quality cuts
described in~\cite{Aamodt:2010pa} were applied to minimize
contamination from secondary charged particles and fake tracks.
The track reconstruction efficiency and contamination
were estimated from HIJING Monte Carlo
simulations~\cite{Wang:1991hta} combined with a GEANT3~\cite{Brun:1994aa} detector model and were found to be independent of
the collision centrality. The reconstruction efficiency increases with transverse momenta from
70\% to 80\% for particles with $0.2<p_{\rm T}<1$~GeV/$c$ and remains
constant at $(80 \pm 5)$\% for $p_{\rm T}>1$~GeV/$c$. The estimated
contamination by secondary charged particles from weak decays and
photon conversions is less than 6\% at $p_{\rm T}=0.2$~GeV/$c$ and falls
below 1\% for $p_{\rm T}>1$~GeV/$c$.
The $p_{\rm T}$ cut-off of 0.2~GeV/$c$ reduces event-by-event biases due to small reconstruction efficiency 
at lower $p_{\rm T}$, while the high $p_{\rm T}$ cut-off of 5~GeV/$c$ reduces the effects of jets on the measured correlations. 
Reconstructed TPC tracks are required to have at least 70 space points (out of a maximum of 159). 
Only tracks with a transverse distance of closest approach to the primary vertex less than 3 mm, both in the longitudinal and transverse directions, are accepted. This reduces the contamination from secondary tracks produced in the detector material, particles from weak decays, etc. Tracks with kinks (i.e. tracks that appear to change direction due to multiple scattering or $K^{\pm}$ decays) were rejected.

@@

\subsection{Centrality Dependence of SC($m$,$n$) and NSC($m$,$n$)}
\label{sec:theory_allorder}
Three different models are compared with the experimental results presented in Sec.~\ref{sec:results}. 
We compare to event-by-event EKRT+viscous hydrodynamic predictions with various parameterizations of the temperature dependence of $\eta/s(T)$; these were shown in Fig.~2 of Ref.~\cite{ALICE:2016kpq}.
In this previous work it was demonstrated that NSC(3,2) is sensitive mainly to the initial conditions, while NSC(4,2) is sensitive to both the initial conditions and the system properties, which is consistent with the predictions from~\cite{Niemi:2012aj}.
%However, the sign of NSC(3,2) is negative in the data in 0--10\% central collisions while it is positive in the models where the anisotropies originate mainly from fluctuations.
%This observation helps us to better understand the fluctuations in initial energy density.
The NSC(4,2) observable shows better sensitivity for different $\eta/s(T)$ parameterizations but the model cannot describe either the centrality dependence or the absolute values. This observed discrepancy between data and theoretical predictions indicates that the current understanding of initial conditions in models of heavy-ion collisions needs to be revisited to further constrain $\eta/s(T)$.
%, considering the difficulties in separating the role of $\eta/s$ from the initial conditions to the final state particle anisotropies~\cite{Romatschke:2007mq,Shen:2011zc}.
The measurement of SC$(m,n)$ and NSC$(m,n)$ can provide new constraints for the detailed modeling of fluctuating initial conditions.
%Better constraints on the initial state conditions will certainly reduce the uncertainties in determining $\eta/s(T)$.

@@

The comparison to the VISH2+1 calculation~\cite{Zhu:2016puf} is shown in Fig.~\ref{fig:Figure_4}.  All calculations with large $\eta/s$ regardless of the initial conditions ($\eta/s=0.2$ for MC-KLN and MC-Glauber initial conditions and $\eta/s=0.16$ for AMPT initial conditions) fail to capture the centrality dependence of the SC($m$,$n$) observables of all orders, shown in the left panels in Fig.~\ref{fig:Figure_4}.
Among the calculations with small $\eta/s$ ($\eta/s=0.08$), the one with the AMPT initial conditions describes the data better than the ones with other initial conditions for all SC($m$,$n$) observables measured, but it cannot describe the data quantitively for most of the centrality ranges.
%Similar to the event-by-event EKRT+viscous hydrodynamic calculations~\cite{Niemi:2015qia}, the sign of the normalized NSC(3,2) in the model calculations in Fig.~\ref{fig:Figure_4} is opposite to that in data in 5--10\% central collisions. 
%NSC(3,2) and NSC(4,3) are not very sensitive to the initial conditions or $\eta/s$ parameterization used in the model, and are not described quantitatively.
The comparisons of the measured NSC(3,2) and NSC(4,3) to VISH2+1 model calculations suggest that they are not very sensitive to the initial conditions or $\eta/s$ parameterization, though the model does not describe them quantitatively.

However, NSC(4,2) is sensitive both to the initial conditions and the $\eta/s$ parameterizations used in the models.
Even though NSC(4,2) favors both AMPT initial conditions with $\eta/s=0.08$ and MC-Glauber initial conditions with $\eta/s=0.20$,
SC(4,2) can only be described by models with smaller $\eta/s$. Hence the calculation with large $\eta/s=0.20$ is ruled out. We conclude that $\eta/s$ should be small and that AMPT initial conditions are favored by the data.
The NSC(5,2) and NSC(5,3) observables are quite sensitive to both the initial conditions and the $\eta/s$ parameterizations.
%Similar to the hydrodynamic calculations mentioned above~\cite{Niemi:2015qia}, the sign of NSC(4,3) in these models is opposite to its sign in the data in 0--10\% central collisions. 
The SC(4,3) results are clearly favored by smaller $\eta/s$ values but NSC(4,3) cannot be described by these models quantitively.

@@

The SC($m$,$n$) and NSC($m$,$n$) observables calculated from AMPT simulations are compared with data in Fig.~\ref{fig:Figure_5}.
For SC(3,2), the calculation with the default AMPT settings is closest to the data, but none of the AMPT configurations can describe the data fully. 
%The same default calculation can describe the sign and magnitude of NSC(3,2) while the hydrodynamic calculations fail to describe either of them in the most central collisions.
%Interestingly, the string melting AMPT configuration cannot reproduce the data, and the resulting correlations are weaker than in data.
The third version based on the string melting configuration without the hadronic rescattering phase is also shown.
The hadronic rescattering stage makes both SC(3,2) and NSC(3,2) stronger in the string melting AMPT model but not enough to describe the data.
Further investigations proved why the default AMPT model can describe NSC(3,2) fairly well but underestimates SC(3,2). By taking the differences in the individual flow harmonics ($v_2$ and $v_3$) between the model and data into account, it was possible to recover the difference in SC(3,2) between the data and the model. The discrepancy in SC(3,2) can be explained by the overestimated individual $v_n$ values as reported in \cite{Adam:2016nfo} in all centrality ranges.

@@

We list here the important findings from the model comparisons to the centrality dependence of SC($m$,$n$) and NSC($m$,$n$):
\begin{enumerate}[(i)]
	\item The NSC(3,2) observable is sensitive mainly to the initial conditions, while the other observables are sensitive to both the initial conditions and the temperature dependence of $\eta/s$.
	\item The correlation strength between $v_3$ and $v_2$ and between $v_4$ and $v_3$ (SC(3,2), SC(4,2), NSC(3,2) and NSC(4,3)) is significantly underestimated in hydrodynamic model calculations in mid-central collisions.
	%\item The sign of NSC(3,2) in 0--10\% central collisions was found to be different in the data and the hydrodynamic model calculations while the default AMPT model can reproduce it.
	\item All the VISH2+1 model calculations with large $\eta/s$ fail to capture the centrality dependence of the correlations regardless of the initial conditions.
	\item Among the VISH2+1 model calculations with small $\eta/s$ ($\eta/s=0.08$), the one with the AMPT initial conditions captures the data qualitatively, but not quantitively for most of the centrality ranges.
	\item The default AMPT model can describe the normalized symmetric cumulants (NSC($m$,$n$)) quantitively for most centralities while the string melting AMPT model fails to describe them.
	%\item The hadronic rescattering in AMPT model makes the correlation strength (NSC) larger.
	\item A hierarchy NSC(5,3) $>$ NSC(4,2) $>$ NSC(5,2) holds for centrality ranges $>~20\%$ within the errors. This hierarchy is reproduced well both by hydrodynamic and AMPT model calculations.

@@

We have found that fluctuations of SC(3,2) and SC(4,3) are anti-correlated in all centralities while fluctuations of SC(4,2), SC(5,2) and SC(5,3) are correlated for all centralities. 
These measurements were compared to various hydrodynamic model calculations with different initial conditions as well as different parameterizations of the temperature dependence of $\eta/s$.
It is found that the different order harmonic correlations have different sensitivities to the initial conditions and the system properties. Therefore they have discriminating power in separating the effects of $\eta/s$  from the initial conditions on the final state particle anisotropies.
%The sign of SC(3,2) in 0--10\% central collisions was found to be different between the data and hydrodynamic model calculations.
%In the most central collisions the anisotropies originate mainly from fluctuations, where the initial ellipsoidal geometry which dominates in mid-central collisions plays little role. This observation might help to understand the details of the fluctuations in the initial stage. 
The comparisons to VISH2+1 calculations show that all the models with large $\eta/s$, regardless of the initial conditions, fail to capture the centrality dependence of higher order correlations. 
Based on the tested model parameters, the data favors small $\eta/s$ and the AMPT initial conditions.
